---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# Contrasts

## Installation

You can install the development version of Contrasts from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("anhsmith/Contrasts")
```

## What does it do?

This package contains functions for converting a factor with $k$ levels into $k-1$ contrasts. The contrasts are a set of numerical variables that encode the levels of a factor.

The contrasts have the following properties:

-   sum to zero
-   the values are inversely weighted by sample sizes of the groups
-   are orthogonal to each other

Also included is a function for making contrasts for a factor `y_in_x` nested in `x`, where a separate set of contrasts are made for the levels of `y_in_x` within each level of `x`.

This code was created so that factors could be coded for inclusion in distance-based linear models (DISTLM) using the software Primer (www.primer-e.co.nz).

## Example

```{r setup, warning = FALSE}
library(Contrasts)
library(dplyr)
```

Make a factor `x` with three levels and uneven sample sizes.

```{r}
x = factor(rep(letters[1:3], times=4:6))
x
```

Make the contrasts.

```{r}
make_weightzero_contrasts(x)
```

Take factor `x`, along with a factor `y_in_x` nested in `x`, and make contrasts for both.

```{r}
y_in_x = factor(paste0(x,c(1,1,2,2,1,1,2,2,3,1,2,2,2,3,3)))

bind_cols(
   x=x, y_in_x=y_in_x,
   make_weightzero_contrasts(x),
   make_weightzero_nested_contrasts(y_in_x, x)
 )
```
